\pkg{webchem}: An \proglang{R} Package to Retrieve Chemical Information from the Web

### 2020-02-11 revision3_2581/paper.tex
Updated table outputs and added tests

In the same manner, we can now query other identifiers from another
source using these CAS numbers (see Figure~\ref{fig:fig1}), like
PubChem
%
\begin{CodeChunk}
\begin{CodeInput}
R> cids <- get_cid(etox_cas)
R> pc_data <- pc_prop(cids, properties = "CanonicalSMILES")
R> pc_smiles <- smiles(pc_data)
\end{CodeInput}
\end{CodeChunk}
%
or ChemSpider
%
\begin{CodeChunk}
\begin{CodeInput}
R> csids <- get_csid(etox_cas, token = token)
R> cs_data <- cs_compinfo(csids, token = token)
R> cs_inchikey <- inchikey(cs_data)
\end{CodeInput}
\end{CodeChunk}
%
Finally, we combine the queried data into one data frame
%
\begin{CodeChunk}
\begin{CodeInput}
R> res <- data.frame(name = subs, cas = etox_cas, smiles = pc_smiles, 
+    cid = pc_data$CID, inchikey = cs_inchikey, csid = cs_data$csid, 
+    stringsAsFactors = FALSE)
\end{CodeInput}
\end{CodeChunk}
%
Note that in order to use the ChemSpider functions, a personal
authentication key (\code{token}) is needed, which can be retrieved
from the ChemSpider web page.  Finally, we obtain a compound table
containing many different identifiers (Table~\ref{tab:comptable}),
allowing easy identification and merging with other data sets, e.g.,
the \code{lc50} data set based on CAS.

@@

\begin{CodeChunk}
\begin{CodeInput}
R> lite <- chebi_lite_entity(lc50_3$cas)
R> lite <- lapply(lite, function(x) x[ which.max(x$searchscore), ]) # highest seachscore
R> lite <- do.call(rbind, lite) # bind to data.frmae
R> lite$cas <- rownames(lite)
R> lite$chebiasciiname[4] <- "tefluthrin" # erronous result
\end{CodeInput}
\end{CodeOutput}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> comp <- chebi_comp_entity(lite$chebiid) # ChEBI ids
R> comp <- do.call(rbind, lapply(comp, `[[`, "parents")) # extract parents table
R> comp$chebiid <- gsub('\\.[0-9]+', '', rownames(comp))
R> comp <- comp[ comp$type == "has role", ] # refine to chemical roles
R> role <- aggregate(chebiName ~ chebiid,
                     data = comp,
                     FUN = function(x) paste0(x, collapse = ', '))
R> merge(lite[ names(lite) %in% c('cas', 'chebiid', 'chebiasciiname') ],
         role,
         by = 'chebiid')
\end{CodeInput}
\end{CodeOutput}

@@

We re-use the above queried ETOX-IDs to obtain further information
from ETOX, namely the MAC-EQS:
% 
\begin{CodeChunk}
\begin{CodeInput}
R> eqs <- etox_targets(ids$etoxid)
R> ids$mac <- sapply(eqs, function(y) {
+    if (length(y) == 1 && is.na(y)) {
+      return(NA) 
+    } else {
+      res <- y$res
+      min(res[res$Country_or_Region == "EEC / EU" & 
+        res$Designation == "MAC-EQS", "Value_Target_LR"])
+    }
+  })
\end{CodeInput}
\end{CodeChunk}
%
Again, the returned information is humongous and we encourage users to study the returned objects and description of the data source.
Here, the column \code{Designation} defines the type of EQS and \code{Value_Target_LR} contains the value.
Unfortunately, we only found MAC-EQS values for 5 substances:
%
\begin{CodeChunk}
\begin{CodeInput}
R> (mac <- with(ids, ids[!is.na(mac) & is.finite(mac), c("etoxid", "query",
+    "mac")]))
\end{CodeInput}
\begin{CodeOutput}
   etoxid       query    mac
4    8397     Atrazin  2.000
5    7240      Benzol 50.000
11   8836     Irgarol  0.016
12   7442 Isoproturon  1.000
29   8756   Terbutryn  0.034
\end{CodeOutput}
\end{CodeChunk}
%
The \code{get_etoxid()} function used to search ETOX-IDs returns also
the original substance name (\code{query}), so that we can easily join
the table with MAC values with the measurements table:
% 
\begin{CodeChunk}
\begin{CodeInput}
R> jagst_eqs <- merge(jagst, mac, by.x = "substance", by.y = "query")
R> head(jagst_eqs)
\end{CodeInput}
\begin{CodeOutput}
  substance       date  value qual etoxid mac
1   Atrazin 2013-09-10 0.0068    =   8397   2
2   Atrazin 2013-10-08 0.0072    =   8397   2
3   Atrazin 2013-03-26 0.0040    =   8397   2
4   Atrazin 2013-04-23 0.0048    =   8397   2
5   Atrazin 2013-11-05 0.0036    =   8397   2
6   Atrazin 2013-07-16 0.0052    =   8397   2
\end{CodeOutput}
\end{CodeChunk}
%
Finally, we can compare the measured value to the MAC, which reveals that there have been no exceedances of these 5 compounds.

>>>>

In the same manner, we can now query other identifiers from another
source using these CAS numbers (see Figure~\ref{fig:fig1}), like
PubChem.
%
\begin{CodeChunk}
\begin{CodeInput}
R> cids <- get_cid(etox_cas)
R> pc_data <- pc_prop(cids, properties = "CanonicalSMILES")
R> pc_smiles <- smiles(pc_data)
\end{CodeInput}
\end{CodeChunk}
%
We can use SMILES to query ChemSpiderIDs and subsequently convert them to InChiKeys.
%
\begin{CodeChunk}
\begin{CodeInput}
R> csids <- get_csid(pc_smiles, from = "smiles")$csid
R> cs_inchikey <- cs_convert(csids, from = "csid", to = "inchikey")
\end{CodeInput}
\end{CodeChunk}
%
Finally, we combine the queried data into one data frame
%
\begin{CodeChunk}
\begin{CodeInput}
R> res <- data.frame(name = subs, cas = etox_cas, smiles = pc_smiles, 
+    cid = pc_data$CID, inchikey = cs_inchikey, csid = cs_data$csid, 
+    stringsAsFactors = FALSE)
\end{CodeInput}
\end{CodeChunk}
%
Note that in order to use the ChemSpider functions, a personal
authentication key is needed, which can be retrieved
from the ChemSpider web page.  Finally, we obtain a compound table
containing many different identifiers (Table~\ref{tab:comptable}),
allowing easy identification and merging with other data sets, e.g.,
the \code{lc50} data set based on CAS.

@@

\begin{CodeChunk}
\begin{CodeInput}
R> lite <- chebi_lite_entity(lc50_3$cas)
R> lite <- lapply(lite, function(x) x[ which.max(x$searchscore), ]) # highest seachscore
R> lite <- do.call(rbind, lite) # bind to data.frmae
R> lite$cas <- rownames(lite)
R> lite$chebiasciiname[4] <- "tefluthrin" # erronous result
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> comp <- chebi_comp_entity(lite$chebiid) # ChEBI ids
R> comp <- do.call(rbind, lapply(comp, `[[`, "parents")) # extract parents table
R> comp$chebiid <- gsub('\\.[0-9]+', '', rownames(comp))
R> comp <- comp[ comp$type == "has role", ] # refine to chemical roles
R> role <- aggregate(chebiName ~ chebiid,
                     data = comp,
                     FUN = function(x) paste0(x, collapse = ', '))
R> merge(lite[ names(lite) %in% c('cas', 'chebiid', 'chebiasciiname') ],
         role,
         by = 'chebiid')
\end{CodeInput}
\end{CodeChunk}

@@

We re-use the above queried ETOX-IDs to obtain further information
from ETOX, namely the MAC-EQS:
% 
\begin{CodeChunk}
\begin{CodeInput}
R> eqs <- etox_targets(ids$etoxid)
R> ids$mac <- sapply(eqs, function(y) {
+    if (length(y) == 1 && is.na(y)) {
+      return(NA) 
+    } else {
+      res <- y$res
+      min(res[res$Country_or_Region == "EEC / EU" & 
+        res$Designation == "MAC-EQS", "Value_Target_LR"])
+    }
+  })
\end{CodeInput}
\end{CodeChunk}
%
Again, the returned information is humongous and we encourage users to study the returned objects and description of the data source.
Here, the column \code{Designation} defines the type of EQS and \code{Value_Target_LR} contains the value.
Unfortunately, we only found MAC-EQS values for 6 substances:
%
\begin{CodeChunk}
\begin{CodeInput}
R> (mac <- with(ids, ids[!is.na(mac) & is.finite(mac), c("etoxid", "query",
+    "mac")]))
\end{CodeInput}
\begin{CodeOutput}
   etoxid       query    mac
4    8397     Atrazin  2.000
5    7240      Benzol 50.000
11   8836     Irgarol  0.016
12   7442 Isoproturon  1.000
38   7571     Simazin  4.000
29   8756   Terbutryn  0.034
\end{CodeOutput}
\end{CodeChunk}
%
The \code{get_etoxid()} function used to search ETOX-IDs returns also
the original substance name (\code{query}), so that we can easily join
the table with MAC values with the measurements table:
% 
\begin{CodeChunk}
\begin{CodeInput}
R> jagst_eqs <- merge(jagst, mac, by.x = "substance", by.y = "query")
R> head(jagst_eqs)
\end{CodeInput}
\begin{CodeOutput}
  substance       date  value qual etoxid mac
1   Atrazin 2013-09-10 0.0068    =   8397   2
2   Atrazin 2013-10-08 0.0072    =   8397   2
3   Atrazin 2013-03-26 0.0040    =   8397   2
4   Atrazin 2013-04-23 0.0048    =   8397   2
5   Atrazin 2013-06-18 0.0048    =   8397   2
6   Atrazin 2013-07-16 0.0052    =   8397   2
\end{CodeOutput}
\end{CodeChunk}
%
Finally, we can compare the measured value to the MAC, which reveals that there have been no exceedances of these 6 compounds.
